Extracting and evaluating educational concept dependencies from Large Language Models

### 2023-03-07 doc/paper.tex
Added graphs to paper

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{neurips_2021}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[colorlinks = true, 
		    linkcolor = blue,
		    urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}       
% hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{natbib}
\usepackage[pdftex]{graphicx}
\usepackage{siunitx} % Required for alignment
\sisetup{
  round-mode          = places, % Rounds numbers
  round-precision     = 2, % to 2 places
}

@@

\section{Introduction}
% Starting broad
% Field of contribution
Large Language Models (LLMs) are trained on immense corpora of text and have proven to build on embedded factual information when performing well in downstream tasks such as question answering. Accessing this knowledge, which is represented by billions of parameters and the network's architecture, has given rise to the research field of Knowledge Extraction from LLMs (Cohen, 2023). Computer-assisted education could greatly benefit from making implicit educational knowledge in LLMs accessible. This creates new perspectives for a subfield of knowledge engineering that has rather stagnated in the previous years.

% What kind of information do we want to extract?
%% Problem of sequencing in Instructional Design:
Concept dependencies are complex but highly relevant educational knowledge. Effective and efficient instruction does not only incorporate what to teach but also how to teach. Guidelines for \textit{instructional sequencing} emphasize the order of instruction. More precisely, prerequisites of educational content should be either known to the student or taught first \citep{morrison2019designing}.
% Definition concepts
Within educational content, Merill (1983) differentiated facts, concepts, principles, rules, procedures, interpersonal skills, attitudes, and their sole recall from their application. To simplify, we will focus on concepts and their relations defined understanding. 
% Concept dependency
% Assumption: information about dependencies between concepts makes instructional sequencing more effective in terms of learning
If one concept is a prerequisite of another, we refer to this relation as a concept dependency. For example, to understand the concept of a derivative, having knowledge about the concept of a function will facilitate or even enable learning. When the dependencies are thought of as directed edges between nodes that represent concepts, a concept dependency graph emerges which is a special type of a knowledge graph \citep{wang2016using}. This graph is also called \textit{concept map} in the field of Learning Sciences.
% TODO: provide a (graphical) example of a small concept map
% Why is is important?
The graph can be used to advance curriculum planning (Yang, 2015), especially for new topics that might not be covered in textbooks or automated assessment (Wang, 2015).

% What is the gap in research that we tackle?
% What is our research question?
% How to extract educational concept dependencies that are implicit and evaluate them using existing knowledge media?
In this work, we tackled the question how to extract this particular knowledge graph and how to evaluate its quality in a scalable manner.

% What can we contribute?
% 1. No benchmark available -> we set up an evaluation framework based on textbooks and Wikipedia
First, there was no classical benchmark available to evaluate the accuracy of extracted information. Therefore, we propose a set of methods to use existing unstructured knowledge sources, namely Wikipedia and textbooks, to create baselines for evaluation. Due to the heuristic character of these methods, we conducted a manual assessment to test their suitability for our purpose. The resulting dataset can be used as a human baseline for further research.

% 2. Emerging field of prompt engineering -> we engineered a simple but effective prompt process to query a LLM
Second, there is no established manner to extract the desired knowledge from the LLM. The emerging field of prompt engineering is currently a vast collection of commands to query LLMs. We propose a method called output refeeding that sequentially queries the language generation model and transforms its answers into a knowledge graph.

@@

% Field: Prompt Engineering for Knowledge Graph creation
% of interest: How can we get internal representations out of a language model?
% Cohen, 2023
% GraphGPT

@@

\subsection{Research design} 
% two baseline methods: textbooks and Wikipedia
% information extraction: one LLM method
% evaluation
% 1. manual inspection via DashBoard
% 2. manual labeling
% 3. convergence between LLM and baselines

\subsection{Baseline extraction}
\subsubsection{Wikipedia}
% Why Wikipedia?
% good coverage according to one of our papers
% 1462 unique articles for book index entries, these were covered in the textbooks (we had to select a subset of concepts to reduce computational effort)
% How did we get the information?
% describe Algorithm to extract dependencies:
%• Assumption: understanding the concept’s definition in Wikipedia requires understanding the concepts in it
%• Operationalization: links to other articles in the first sentence of article
%• Filtered out links to fields (“Mathematics”), persons (“Gabriel Cramer), and concepts that linked even earlier to the concept of interest

\subsubsection{Textbooks}
% Types of textbooks (why linear algebra?)

% Amount of data
% 10 books on Linear Algebra
% 584 pages on average (652,297 characters)
% 382 number of index entries (median)

% Steps of processing
% preprocessing
% entity recognition and disambigation
% Wikifier: identify entity mentions using link targets and PageRank
For preprocessing the textbooks we used Wikifier \citep{brank2017annotating}.
% Wikisearch: query Wikipedia search engine for keyword (and field) and rank results based on Levenshtein distance

@@

\subsection{Manual inspection via Dashboard}
%

@@

@@

@@

% What are potential flaws?
% disambiguation already includes noise
% we only evaluated up to two levels of the graph, not the graph as a whole. different level of abstractions in extracted dependencies
% is there even something like a gold standard? some educational reflection about different didactic approaches and more depth than just understanding the definition of a topic
% we are still using the language generator, so it is just another downstream task, maybe it is more of a philosophical question what is implicit/explicit knowledge

% Q: Outlook and Future research ideas:
% advancements in LLMs that are interesting for us
% use cases in education

@@

>>>>

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{neurips_2021}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[colorlinks = true, 
		    linkcolor = blue,
		    urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}       
% hyperlinks
\usepackage{float}
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{natbib}
\usepackage[pdftex]{graphicx}
\usepackage{siunitx} % Required for alignment
\sisetup{
  round-mode          = places, % Rounds numbers
  round-precision     = 2, % to 2 places
}

@@

\section{Introduction}
% Starting broad
% Field of contribution
Large Language Models (LLMs) are trained on immense corpora of text and have proven to build on embedded factual information when performing well in downstream tasks such as question answering. Accessing this knowledge, which is represented by billions of parameters and the network's architecture, has given rise to the research field of Knowledge Extraction from LLMs (\cite{cohen2023crawling}), which rests on the assumption that the language models can help retrieve information on the relation between entities.

This work tries to contribute to the research field and to answer the question of whether general LLMs are able to provide a valid dependency structure of educational concepts. To find the answer to the research question, we are going to solve two main subtasks: 1) extraction of implicit educational concept dependencies, and 2) evaluation of the retrieved dependencies or the resulting graphs by existing knowledge media.

Apart from the field of knowledge extraction from LLMs, which is a subfield of computational linguistics, our work may contribute to applied education sciences by improving computer-assisted education and creating new perspectives for a subfield of knowledge engineering that has rather stagnated in the previous years.

% What kind of information do we want to extract?
%% Problem of sequencing in Instructional Design:
In particular, effective and efficient instruction does not only incorporate what to teach but also how to teach, especially the order of instruction.  It is proven that \textit{instructional sequencing} enhances the process of learning when prerequisites of educational content are known to the student or taught first \citep{morrison2019designing}.
% Definition concepts
Within educational content, Merill (1983) differentiated facts, concepts, principles, rules, procedures, interpersonal skills, attitudes, and their sole recall from their application. We will focus on concepts and their relations defined understanding. 
% Concept dependency
% Assumption: information about dependencies between concepts makes instructional sequencing more effective in terms of learning

To be more precise, if one concept is a prerequisite of another, we refer to this relation as a concept dependency. For example, to understand the concept of a derivative, having knowledge about the concept of a function will facilitate or even enable learning. When the dependencies are thought of as directed edges between nodes that represent concepts, a concept dependency graph emerges which is a special type of a knowledge graph \citep{wang2016using}. This graph is also called \textit{concept map} in the field of Learning Sciences.
% TODO: provide a (graphical) example of a small concept map
% Why is is important?
The graph can be used to advance curriculum planning (\cite{yang2015concept}), especially for new topics that might not be covered in textbooks or automated assessment (Wang, 2015).

% What is the gap in research that we tackle?
% What is our research question?
% How to extract educational concept dependencies that are implicit and evaluate them using existing knowledge media?
%In this work, we tackled the question how to extract this particular knowledge graph and how to evaluate its quality in a scalable manner.

% What can we contribute?
% 1. No benchmark available -> we set up an evaluation framework based on textbooks and Wikipedia
To conclude, our main contributions are the following.
First, there is little research on how to evaluate the precision of extracted concept dependencies. Therefore, we propose a set of methods to use existing unstructured knowledge sources, namely Wikipedia and textbooks, to create baselines for evaluation. Due to the heuristic character of these methods, we conducted a manual assessment to test their suitability for our purpose. The resulting dataset can be used as a baseline for further research.

% 2. Emerging field of prompt engineering -> we engineered a simple but effective prompt process to query a LLM
Second, the emerging field of prompt engineering provides a constellation of commands to query LLMs, the majority of which are experimental and cannot be considered fully reliable. We propose a method called \textit{output refeeding} that allows mining the educational concept dependencies by sequentially querying the language generation model and transforming its answers into a knowledge graph.

@@

% Field: Prompt Engineering for Knowledge Graph creation
% of interest: How can we get internal representations out of a language model?
% Cohen, 2023

Multiple approaches of prompt engineering for graph creation have been recently developed. Few-shot prompting proved successful (\cite{cohen2023crawling})

% GraphGPT

@@

\subsection{Research design}

Our research lies at the intersection of education sciences and computational linguistics that define the choice of methodology. On the one hand, we prompt a state-of-the-art large language model to build entity dependency graphs and evaluate its performance on the concept relations retrieved from Wikipedia and textbooks. On the other hand, we should ensure that the output is sound in terms of educational value, which requires additional qualitative assessment.

Since no ground-truth knowledge graph exists and cannot be achieved in the field of education, we choose textbooks and Wikipedia as the baseline to be able to estimate the precision of the dependencies mined from the LLM. Each information extraction method is not fully independent of the other, in particular, entity recognition and concept disambiguation steps in the procedure for LLM and textbooks depend on Wikipedia. The evaluation was performed via 1) inspecting the output in the interactive dashboard, 2) manual labeling, and 3) convergence between the LLM and two baselines metric.

% two baseline methods: textbooks and Wikipedia
% information extraction: one LLM method
% evaluation
% 1. manual inspection via DashBoard
% 2. manual labeling
% 3. convergence between LLM and baselines

\subsection{Baseline extraction}
\subsubsection{Wikipedia}
% Why Wikipedia? % I asked ChatGPT :)

Wikipedia is a valuable resource that has a good structure of concepts. One of the key factors that contribute to it is the use of a hierarchical structure. Articles are categorized into broader topics, and those topics are further divided into subtopics. This structure helps readers to navigate through the information in a logical and organized manner, allowing them to quickly find the information they need. Another factor is the use of hyperlinks to connect related articles. This allows readers to move seamlessly between articles and explore related concepts without having to leave the site or perform additional searches. Wikipedia also has a strong editorial process that ensures that articles are well-researched, well-written, and up-to-date. This process helps to maintain the quality and accuracy of the information presented on the site, which in turn helps to reinforce the logical structure of the concepts presented.

% good coverage according to one of our papers
% 1462 unique articles for book index entries, these were covered in the textbooks (we had to select a subset of concepts to reduce computational effort)
% How did we get the information?
% describe Algorithm to extract dependencies:
%• Assumption: understanding the concept’s definition in Wikipedia requires understanding the concepts in it
%• Operationalization: links to other articles in the first sentence of article
%• Filtered out links to fields (“Mathematics”), persons (“Gabriel Cramer), and concepts that linked even earlier to the concept of interest

\subsubsection{Textbooks}
% Types of textbooks (why linear algebra?) % I asked ChatGPT :)

We chose textbooks on linear algebra as they tend to have a good structure and order compared to other textbooks because linear algebra is a well-established and mature field of mathematics with a clear and logical progression of concepts. Moreover, linear algebra is used extensively in other fields, such as engineering, physics, and computer science. Therefore, many textbooks on linear algebra are written with the needs of these fields in mind, which helps to structure the content of the textbook in a clear and practical manner. Finally, the concepts in linear algebra are interrelated, and a strong understanding of earlier topics is often necessary to understand later topics.

We used 10 books that are free and available online. 
% Amount of data
% 10 books on Linear Algebra
% 584 pages on average (652,297 characters)
% 382 number of index entries (median)

% Steps of processing
% preprocessing
% entity recognition and disambigation
% Wikifier: identify entity mentions using link targets and PageRank
First, the textbooks were converted from pdf format to raw texts and went through basic preprocessing. Then, Wikifier \citep{brank2017annotating} was exploited for entity recognition and disambiguation.

% Wikisearch: query Wikipedia search engine for keyword (and field) and rank results based on Levenshtein distance

@@

\begin{figure}
    \centering
    \includegraphics[width=.95\textwidth]{img/dash_example.png}
    \caption{A screenshot of the intercative application.}
    \label{fig:dash_example}
\end{figure}

\subsection{Manual inspection via Dashboard}
We developed a Dashboard via graphing library for Python Plotly\footnote{https://plotly.com/python/} (Fig.\ref{fig:dash_example}). The application is devised so that one could examine the resulting graph itself in a convenient environment. It allows plotting the concept dependencies graphs for three knowledge sources or types of dependency mining methods (LLM, Wikipedia, and textbooks) for different levels of dependencies.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=.95\textwidth]{img/rating.pdf}
    \caption{Rating.}
    \label{fig:rating}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.4\textwidth]{img/kappa.pdf}
    \caption{Interrater reliability, kappa coefficient.}
    \label{fig:kappa}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=.95\textwidth]{img/comp_direct_deps.pdf}
    \caption{Convergence metrics. Number of common direct dependencies between sources.}
    \label{fig:comp_direct_deps}
\end{figure}

@@

% What are potential flaws?

% disambiguation already includes noise
% error propagation

% we only evaluated up to two levels of the graph, not the graph as a whole. different level of abstractions in extracted dependencies
% is there even something like a gold standard? some educational reflection about different didactic approaches and more depth than just understanding the definition of a topic
% Cohen et al (2023) think that there is no gold standard, ground truth.

% we are still using the language generator, so it is just another downstream task, maybe it is more of a philosophical question what is implicit/explicit knowledge

% + "I don't know" option in the prompt engineering part
% + Google search

% Q: Outlook and Future research ideas:

As shown, current LLMs are not sufficiently precise in creating concept maps, which confirms previous research \cite{hwang2021comet}. This could be improved by training the pre-trained models on the knowledge graph data mined from Wikipedia or any other reliable sources of concept dependencies in a specific subject of instruction \cite{west2022symbolic}.
% advancements in LLMs that are interesting for us
% use cases in education

@@

\newpage

\section{Appendix}

\begin{figure}[H]
    \centering
    \includegraphics[width=.95\textwidth]{img/comp_second_deps.pdf}
    \caption{Convergence metrics. Number of common direct and second-level dependencies between sources.}
    \label{fig:comp_second_deps}
\end{figure}
